The Birth of a Galaxy: Population III Progenitors and Metal-enriched Stellar Populations

### 2010-10-13 ms.tex
Wrote Pop III section, started on Pop II section.

The first (Pop III) stars are metal-free and have a large
characteristic mass and suppressed fragmentation in its protostellar
collapse \citep{Abel00, ABN02, Bromm02_P3, Yoshida03, OShea07a}.  A
fraction of these stars enrich the surrounding intergalactic medium
(IGM) when they go supernova, which can happen in stars $\lsim40~\Ms$
in Type II supernovae (SNe) or in stars roughly between 140 \Ms~and
260 \Ms~in pair-instability SNe \citep[PISNe;][]{2002ApJ...567..532H}.
The host halo and the neighboring halos are then enriched with this
ejecta.  There exists a critical metallicity that is $\sim 10^{-6}
Z_\odot$ if dust cooling is efficient \citep{Omukai05,
  Schneider06_Frag, clark08} and $\sim 10^{-3.5} Z_\odot$ otherwise
\citep{Bromm01, 2009ApJ...691..441S}, where the gas can cool rapidly,
lowering its Jeans mass.  An intermediate characteristic mass of
$\sim10~\Ms$ can be occur if the gas cooling is suppressed to the
cosmic microwave background (CMB) temperature \citep{Larson98,
  Tumlinson07_IMF, 2009ApJ...691..441S}.  The resulting Population II
star cluster will thus have a lower characteristic stellar mass than
its metal-free progenitors.  In the local universe, stars in the Milky
Way halo and dwarf spheroidal (dSph) galaxies may be connected to
these first stellar clusters.  In the halo of the Milky Way, the most
metal-poor stars have a metallicity floor of [Z/H] = -4
\citep{Beers05}, and in dwarf spheroidal (dSph) galaxies this floor is
a factor of 10 higher \citep{Tolstoy04, Helmi06}.  \textbf{Check for
  more recent references.}

@@

We use the adaptive mesh refinement (AMR) code
\enzo~\citep{BryanNorman1997, OShea2004} that employs our
implementation of adaptive ray tracing for radiation transport
\citep{Wise10}.  To resolve minihalos with at least 100 dark matter
(DM) particles and follow the formation of the first generation of
dwarf galaxies, we use a simulation box of 1 Mpc that has a resolution
of $256^3$.  This gives us a DM mass resolution of 1840 \Ms.  We
refine the grid on baryon overdensities of $3 \times 2^{-0.2l}$, where
$l$ is the AMR level, resulting in a super-Lagrangian behavior.  We
also refine on a DM overdensity of $3 \times 2^l$ and always resolve
the local Jeans length by at least 4 cells.  The latter criterion
avoids any artifical fragmentation during gaseous collapses
\citep{Truelove97}.  We initialize the simulation with \textsl{grafic}
\citep{Bertschinger01} at $z = 130$ and use the cosmological
parameters from the 7-year WMAP data \citep{WMAP7}: $\Omega_M =
0.266$, $\Omega_\Lambda = 0.734$, $\Omega_b = 0.0449$, $h = 0.71$,
$\sigma_8 = 0.81$, and $n = 0.963$ with the variables having their
usual definitions.

@@

We distinguish Pop II and Pop III star formation by the metallicity of
the cold ($T < 10^3$ K) star forming gas in the molecular clouds.  Pop
II stars are formed if [Z/H] $> -4$, and Pop III stars are formed
otherwise.  We do not consider Pop III.2 stars and intermediate mass
stars from CMB-limited cooling.

@@

For Pop III stars, we use the same star formation model as
\citet{Abel07} and \citet{Wise08_Gal} where each star particle
represents a single star.  The critical overdensity in which a star
forms is $5 \times 10^5$.  Instead of using a fixed stellar mass, we
randomly sampled from an IMF with a functional form of
%
\begin{equation}
\label{eqn:imf}
f(M)dM = M^{-1.3} \exp\left[-\left(\frac{M_{\rm char}}{M}\right)^{1.6}\right]
\end{equation}
to determine the stellar mass.  Above $M_{\rm char}$, it behaves as a
Salpeter IMF but is exponentially cutoff below that mass
\citep{Chabrier03, Clark09}.  For reproducibility, we initialize both
simulations with the same random seed and record the number of times
the random number generator \citep[Mersenne twister;][]{MTwister} has
been called for use when restarting the simulations.

@@

\begin{figure*}
  \plotone{f3}
  \caption{\label{fig:projections} Density-weighted projections of gas
    density (top), temperature (middle), and metallicity (bottom).
    The left column shows the entire simulation volume, where the
    center and right columns focus on the intense and quiet halos,
    respectively.  The metallicity projections are a composite picture
    of metals originating from Pop III (red) and Pop II (blue) stars.}
\end{figure*}

@@

\subsection{Halo evolution}

@@

At $z \sim 8$, the metallicity of the intense halo dramatically
self-enriches itself by a factor of 30 to $[\bar{Z}_2/H] = 1.5$ during
a starburst.  Because this halo is located in a large-scale
overdensity, the ejecta falls back into the halo after reaching
distances up to 20 comoving kpc, in turn keeping the halo gas
metallicity high because the inflows are relatively metal-rich
themselves.  Interestingly the first few Pop II star clusters have
[Z/H] between --1 and --2 and dominate the average stellar metallicity
at $z > 8$.  We further detail the formation of these clusters in
\S\ref{sec:pop}.  After the $z=8$ starburst, the average stellar
metallicity follows the average gas metallicity within 0.1 dex.

@@

\li Report on the Pop III stars forming in halo progenitors

The most massive progenitor of the quiet halo interestingly never
hosts a Pop III star.  Instead a nearby halo forms a Pop III star,
which (randomly) produces a PISNe at $z=16$.  The associated blastwave
overruns the most massive progenitor, and the dense core survives this
event and is enriched by this PISN.

@@

\li Present results on the star formation history and the physical
reasons for the features seen in the SFH and metallicities.

\li Present the stellar metallicity distribution at the final time

>>>>

The first (Pop III) stars are metal-free and have a large
characteristic mass and suppressed fragmentation in its protostellar
collapse \citep{Abel00, ABN02, Bromm02_P3, Yoshida03, OShea07a}.  A
fraction of these stars enrich the surrounding intergalactic medium
(IGM) when they go supernova, which can happen in stars $\lsim40~\Ms$
in Type II supernovae (SNe) or in stars roughly between 140 \Ms~and
260 \Ms~in pair-instability SNe \citep[PISNe;][]{2002ApJ...567..532H}.
The host halo and the neighboring halos are then enriched with this
ejecta.  There exists a critical metallicity that is $\sim 10^{-6}
Z_\odot$ if dust cooling is efficient \citep{Omukai05,
  Schneider06_Frag, clark08} and $\sim 10^{-3.5} Z_\odot$ otherwise
\citep{Bromm01, 2009ApJ...691..441S}, where the gas can cool rapidly,
lowering its Jeans mass.  An intermediate characteristic mass of
$\sim10~\Ms$ can be occur if the gas cooling is suppressed to the
cosmic microwave background (CMB) temperature \citep{Larson98,
  Tumlinson07_IMF, 2009ApJ...691..441S}.  The resulting Population II
star cluster will thus have a lower characteristic stellar mass than
its metal-free progenitors.  In the local universe, stars in the Milky
Way halo and dwarf spheroidal (dSph) galaxies may be connected to
these first stellar clusters.  In the halo of the Milky Way, the most
metal-poor stars have a metallicity floor of [Z/H] = --4
\citep{Beers05}, and in dwarf spheroidal (dSph) galaxies this floor is
a factor of 10 higher \citep{Tolstoy04, Helmi06}.  \textbf{Check for
  more recent references.}

@@

We use the adaptive mesh refinement (AMR) code
\enzo~\citep{BryanNorman1997, OShea2004} that employs our
implementation of adaptive ray tracing for radiation transport
\citep{Wise10}.  To resolve minihalos with at least 100 dark matter
(DM) particles and follow the formation of the first generation of
dwarf galaxies, we use a simulation box of 1 Mpc that has a resolution
of $256^3$.  This gives us a DM mass resolution of 1840 \Ms.  We
refine the grid on baryon overdensities of $3 \times 2^{-0.2l}$, where
$l$ is the AMR level, resulting in a super-Lagrangian behavior.  We
also refine on a DM overdensity of $3 \times 2^l$ and always resolve
the local Jeans length by at least 4 cells.  The latter criterion
avoids any artifical fragmentation during gaseous collapses
\citep{Truelove97}.  We initialize the simulation with \textsl{grafic}
\citep{Bertschinger01} at $z = 130$ and use the cosmological
parameters from the 7-year WMAP data \citep{WMAP7}: $\Omega_M =
0.266$, $\Omega_\Lambda = 0.734$, $\Omega_b = 0.0449$, $h = 0.71$,
$\sigma_8 = 0.81$, and $n = 0.963$ with the variables having their
usual definitions.  The size of the box limits the simulation to an
ending point of $z=7$.

@@

We distinguish Pop II and Pop III star formation by the metallicity of
the densest cell in the molecular cloud.  Pop II stars are formed if
[Z/H] $> -4$, and Pop III stars are formed otherwise.  We do not
consider Pop III.2 stars and intermediate mass stars from CMB-limited
cooling.

@@

For Pop III stars, we use the same star formation model as
\citet{Abel07} and \citet{Wise08_Gal} where each star particle
represents a single star.  The critical overdensity in which a star
forms is $5 \times 10^5$.  Instead of using a fixed stellar mass, we
randomly sampled from an IMF with a functional form of
%
\begin{equation}
\label{eqn:imf}
f(M)dM = M^{-1.3} \exp\left[-\left(\frac{M_{\rm char}}{M}\right)^{1.6}\right]
\end{equation}
to determine the stellar mass.  Above $M_{\rm char}$, it behaves as a
Salpeter IMF but is exponentially cutoff below that mass
\citep{Chabrier03, Clark09}.  For reproducibility, we record the
number of times the random number generator \citep[Mersenne
  twister;][]{MTwister} has been called for use when restarting the
simulations.

@@

\begin{figure*}
  \plotone{f3}
  \caption{\label{fig:projections} Density-weighted projections of gas
    density (top), temperature (middle), and metallicity (bottom) at
    $z=7$.  The left column shows the entire simulation volume, where
    the center and right columns focus on the intense and quiet halos,
    respectively.  The metallicity projections are a composite picture
    of metals originating from Pop III (red) and Pop II (blue) stars.}
\end{figure*}

@@

\subsection{Halo evolution}
\label{sec:halo}

@@

At $z \sim 8$, the metallicity of the intense halo dramatically
self-enriches itself by a factor of 30 to $[\bar{Z}_2/H] = -1.5$
during a starburst.  Because this halo is located in a large-scale
overdensity, the ejecta falls back into the halo after reaching
distances up to 20 comoving kpc, in turn keeping the halo gas
metallicity high because the inflows are relatively metal-rich
themselves.  Interestingly the first few Pop II star clusters have
[Z/H] between --1 and --2 and dominate the average stellar metallicity
at $z > 8$.  We further detail the formation of these clusters in
\S\ref{sec:pop}.  After the $z=8$ starburst, the average stellar
metallicity follows the average gas metallicity within 0.1 dex.

@@

\textit{Quiet halo}.--- The most massive progenitor interestingly
never hosts a Pop III star.  Instead a nearby halo forms a Pop III
star, which (randomly) produces a PISNe at $z=15.6$.  The associated
blastwave overruns the most massive progenitor, and the dense core
survives this event and is enriched by this PISN, triggering the
transition to Pop II star formation.  Other progenitors host three Pop
III stars, forming at $z = 15.4, 14.2, 13.8$, with the latter
producing a PISN.  Metal enrichment from these two PISNe and latter
intrinsic Pop II SF quench all Pop III star formation in this halo.

\textit{Intense halo}.--- The progenitors host a total of 56 Pop III
stars with 21 producing PISNe.  The first forms in a $6 \times 10^5
\Ms$ halo at $z=19$.  Pop III stars form on a regular interval in the
halo's progenitors until $z=9$ when most of these halos enter the
metal-rich bubble surrounding the intense halo.  There are two Pop III
stars forming at $z=7.7$ with [Z/H] = --5.7, whose molecular clouds
were enriched by external blastwaves that only partially enriched the
dense regions \citep{Cen08}.  The most massive progenitors of the
intense halo cease to form Pop III stars at $z<12$ because of nearby
or intrinsic PISNe.

@@

%% \li Present results on the star formation history and the physical
%% reasons for the features seen in the SFH and metallicities.
%%
%% \li Present the stellar metallicity distribution at the final time

Figure \ref{fig:pops} shows the star formation history of both halos.
From this plot, the metallicity distribution, star formation rates,
and metallicity evolution and spread at a given time can all be
extracted from these data.

\textit{Quiet halo}.--- 
A nearby PISNe provides a metallicity floor of [Z/H] = --2.8 at which
metallicity the first Pop II stars form.  The stellar metallicity
evolution exhibits what is expected from a relatively isolated system
with the stellar feedback steadily enriching the ISM, resulting in a
correlation between stellar age and metallicity.  After $z=10$ the
metallicities plateau at [Z/H] = --2.1 for reasons discussed in
\S\ref{sec:halo}.  Around $z=7.5$, a 25:1 minor merger occurs, and the
gas inside the satellite halo is compressed, triggering metal-poor,
[Z/H] = --3.2, SF during its nearest approach.  This halo infalls
through a filament, where most of metal enrichment in the quiet halo
occurs in bi-polar flows perpendicular to the galaxy disk and
filament.  The filament and thus this satellite halo remains
relatively metal-poor.  The bi-polar outflows and metal-poor filaments
are seen in the metallicity projections in Figure
\ref{fig:projections}.

\textit{Intense halo}.---
